---
title: "Global FoodOmics reference sample interpretation"
authors: Julia M. Gauglitz, Candace Williams
output: html_document
---

#########################################
Figure 2 results and figure generation
#########################################

Generation of results from molecular networking output.
Input data - load one dataframe at a time and run notebook cells. Will generate values in-line, which were transferred to a data table manually.
```{r}
#Read in GNPS output file "View All Clusters With IDs" (.tsv)

GNPS_network_data <- read.delim(file = "../data/Rady CMC (G1) - formulas (G3) - FoodOmics 3500 (no formula) (G4) FDR 0.01/METABOLOMICS-SNETS-V2-df244238-view_all_clusters_withID_beta-main.tsv", sep="\t")

GNPS_network_data <- read.delim(file = "../data/ONR Fecal match - ONR Plasma match - ONR Food - FoodOmics 3500 (no ONR) FDR 0.01/METABOLOMICS-SNETS-V2-9a90bd12-view_all_clusters_withID_beta-main.tsv", sep="\t")

GNPS_network_data <- read.delim(file = "../data/ONR Fecal - ONR Plasma - ONR Food - FoodOmics 3500 (no ONR) FDR 0.01/METABOLOMICS-SNETS-V2-e0bf255b-view_all_clusters_withID_beta-main.tsv", sep="\t")

GNPS_network_data <- read.delim(file = "../data/ONR Fecal - ONR Plasma - CIAO Food - FoodOmics 3500 (no CIAO) FDR 0.01/METABOLOMICS-SNETS-V2-6fe43476-view_all_clusters_withID_beta-main.tsv", sep="\t")

GNPS_network_data <- read.delim(file = "../data/MSV000082431 IBD - FoodOmics 3500 FDR 0.1/METABOLOMICS-SNETS-V2-ec08eed8-view_all_clusters_withID_beta-main.tsv", sep="\t")

GNPS_network_data <- read.delim(file = "../data/GFOP3500 and MSV000082221 IBD SEED/METABOLOMICS-SNETS-V2-907f2e0b-view_all_clusters_withID_beta-main.tsv", sep="\t")

GNPS_network_data <- read.delim(file = "../data/FoodOmics 3500 FDR 0.1/METABOLOMICS-SNETS-V2-d5adba7f-view_all_clusters_withID_beta-main.tsv", sep="\t")

GNPS_network_data <- read.delim(file = "../data/Danone - all - FoodOmics 3500 FDR 0.1/METABOLOMICS-SNETS-V2-5cca39e0-view_all_clusters_withID_beta-main.tsv", sep="\t")

GNPS_network_data <- read.delim(file = "../data/CIAO Fecal - plasma - ONR food - Foodomics 3500 (no ONR food) FDR 0.01/METABOLOMICS-SNETS-V2-981c9a7d-view_all_clusters_withID_beta-main.tsv", sep="\t")

GNPS_network_data <- read.delim(file = "../data/CIAO Fecal - plasma - food - Foodomics 3500 (no CIAO food) FDR 0.01/METABOLOMICS-SNETS-V2-8895b6e3-view_all_clusters_withID_beta-main.tsv", sep="\t")

GNPS_network_data <- read.delim(file = "../data/GFOP3500/AGP - FoodOmics 3500 FDR 0.1/METABOLOMICS-SNETS-V2-4884483b-view_all_clusters_withID_beta-main.tsv", sep="\t")

GNPS_network_data <- read.delim(file = "../data/GFOP3500/Malawi stool - GFOP3500 overlap/METABOLOMICS-SNETS-V2-831366ba-view_all_clusters_withID_beta-main.tsv", sep="\t")

GNPS_network_data <- read.delim(file = "../data/Vaccine response - FoodOmics 3500 FDR 0.01/METABOLOMICS-SNETS-V2-08e9b9e0-view_all_clusters_withID_beta-main.tsv", sep="\t")

GNPS_network_data <- read.delim(file = "../data/Vaccine response - breastmilk - FoodOmics 3500 FDR 0.01/METABOLOMICS-SNETS-V2-08de24ed-view_all_clusters_withID_beta-main.tsv", sep="\t")

GNPS_network_data <- read.delim(file = "../data/MSV000084908 serum - GFOP3500 FDR 0.01/METABOLOMICS-SNETS-V2-55bef022-view_all_clusters_withID_beta-main.tsv", sep="\t")

GNPS_network_data <- read.delim(file = "../data/MSV000082220 SEED Grant IBD Biopsy - GFOP3500/METABOLOMICS-SNETS-V2-a83a279d-view_all_clusters_withID_beta-main.tsv", sep="\t")

GNPS_network_data <- read.delim(file = "../data/MSV000079115 IBD Fecal Sample Solvent Test - GFOP3500/METABOLOMICS-SNETS-V2-fad74693-view_all_clusters_withID_beta-main.tsv", sep="\t")

GNPS_network_data <- read.delim(file = "../data/MSV000084908 fecal - GFOP3500 FDR 0.01/METABOLOMICS-SNETS-V2-709fbdea-view_all_clusters_withID_beta-main.tsv", sep="\t")

GNPS_network_data <- read.delim(file = "../data/MSV000084775 IBD - GFOP3500/METABOLOMICS-SNETS-V2-07f85565-view_all_clusters_withID_beta-main.tsv", sep="\t")

GNPS_network_data <- read.delim(file = "../data/MSV000079777 IBD Biobank Stool Samples - GFOP3500/METABOLOMICS-SNETS-V2-a79fbd4c-view_all_clusters_withID_beta-main.tsv", sep="\t")

GNPS_network_data <- read.delim(file = "../data/Doty fecal - breastmilk - GFOP3500/METABOLOMICS-SNETS-V2-a7b22246-view_all_clusters_withID_beta-main.tsv", sep="\t")

GNPS_network_data <- read.delim(file = "../data/Doty oral - breastmilk - GFOP3500/METABOLOMICS-SNETS-V2-c756a9df-view_all_clusters_withID_beta-main.tsv", sep="\t")

GNPS_network_data <- read.delim(file = "../data/Doty skin - breastmilk - GFOP3500/METABOLOMICS-SNETS-V2-4a322e64-view_all_clusters_withID_beta-main.tsv", sep="\t")

GNPS_network_data <- read.delim(file = "../data/Vaccine response - breastmilk_all - FoodOmics 3500 FDR 0.01/METABOLOMICS-SNETS-V2-2a51ab24-view_all_clusters_withID_beta-main.tsv", sep="\t")

GNPS_network_data <- read.delim(file = "../data/RA fecal - plasma - food - FoodOmics 3500 FDR 0.01 tol 0.01 min 2/METABOLOMICS-SNETS-V2-0794151f-view_all_clusters_withID_beta-main.tsv", sep="\t")

GNPS_network_data <- read.delim(file = "../data/AD Plasma - CSF - FoodOmics 3500 FDR 0.01 tol 0.01 2 mincluster/METABOLOMICS-SNETS-V2-aac78e9d-view_all_clusters_withID_beta-main.tsv", sep="\t")

GNPS_network_data <- read.delim(file = "../data/Saliva SEED - GFOP3500/METABOLOMICS-SNETS-V2-6dd6e5b1-view_all_clusters_withID_beta-main.tsv", sep="\t")

GNPS_network_data <- read.delim(file = "../data/Urine - legume - FoodOmics 3500 (no legumes) FDR 0.01/METABOLOMICS-SNETS-V2-93ba727a-view_all_clusters_withID_beta-main.tsv", sep="\t")

GNPS_network_data <- read.delim(file = "../data/40e855040632405b9c9f9e9e9d5c5bfb/METABOLOMICS-SNETS-V2-40e85504-view_all_clusters_withID_beta-main.tsv", sep="\t")

#covid-19
GNPS_network_data <- read.delim(file = "../data/MSV000085505 covid-19 - MSV000085537 - GFOP3500 FDR 0.1 current release/METABOLOMICS-SNETS-V2-9cbcb6b4-view_all_clusters_withID_beta-main.tsv", sep="\t")

```

Create data tables based on group occurrence (for datasets with up to 4 groups)
Human data was contained in G1 and G2, SSF in G3, and GFOP3500 dataset in G4
```{r}

G1_occurrence <- GNPS_network_data[GNPS_network_data$G1 > 0, ]
G2_occurrence <- GNPS_network_data[GNPS_network_data$G2 > 0, ]
G3_occurrence <- GNPS_network_data[GNPS_network_data$G3 > 0, ]
G4_occurrence <- GNPS_network_data[GNPS_network_data$G4 > 0, ]

#load packages required for dataset modulation
library(tidyr)
library(dplyr)

#fecal-food overlaps
G1G3_occurrence <- G1_occurrence %>% 
  mutate(overlap_binary = ifelse(G3 !=0 & G4 ==0, 1, 0))

G1G4_occurrence <- G1_occurrence %>% 
  mutate(overlap_binary = ifelse(G4 !=0 & G3 ==0, 1, 0))

G1G3G4_occurrence <- G1_occurrence %>% 
  mutate(overlap_binary = ifelse(G3 !=0 & G4 !=0, 1, 0))

#plasma-food overlaps
G2G3_occurrence <- G2_occurrence %>% 
  mutate(overlap_binary = ifelse(G3 !=0 & G4 ==0, 1, 0))

G2G4_occurrence <- G2_occurrence %>% 
  mutate(overlap_binary = ifelse(G4 !=0 & G3 ==0, 1, 0))

G2G3G4_occurrence <- G2_occurrence %>% 
  mutate(overlap_binary = ifelse(G3 !=0 & G4 !=0, 1, 0))

```

Calculates 4 values: 
% annotation in original dataset
% spectra explained by overlap with G3 and annotations
% spectra explained by overlap with G4 and annotations (these two numbers do not sum, because some compounds are observed in both G3 and G4)
% spectra explained by overlap with G3 and G4 -- total interpreted
# output validated by manual calculation with cytoscape network
```{r}
percent_annotation_overlap <- function(occurrence_table_d,G1G3_occurrence,G1G4_occurrence,G1G3G4_occurrence) {
  num_annot <-sum(occurrence_table_d$LibraryID != "N/A") #number of annotations in occurrence table
  dataframe_dim <- data.frame(dim(occurrence_table_d)) 
  total_metabolites <- dataframe_dim[1,1]                               
  percent <- num_annot / total_metabolites
 
  G3_added_info <- sum(G1G3_occurrence$overlap_binary == 1) -
  sum(G1G3_occurrence$overlap_binary == 1 & G1G3_occurrence$LibraryID != "N/A")
  
  G4_added_info <- sum(G1G4_occurrence$overlap_binary == 1) - 
  sum(G1G4_occurrence$overlap_binary == 1 & G1G4_occurrence$LibraryID != "N/A")
    
  G3G4_added_info <- sum(G1G3G4_occurrence$overlap_binary == 1) - 
  sum(G1G3G4_occurrence$overlap_binary == 1 & G1G3G4_occurrence$LibraryID != "N/A")
  
  total_metabolites_G3inclusive <- num_annot + G3_added_info + G3G4_added_info
  total_metabolites_G4inclusive <- num_annot + G4_added_info + G3G4_added_info
  total_metabolites_food_inclusive <- num_annot + G3_added_info + G4_added_info + G3G4_added_info
  
  perc_increase_G3 <- (total_metabolites_G3inclusive/total_metabolites)*100
  perc_increase_G4 <- (total_metabolites_G4inclusive/total_metabolites)*100
  perc_increase_allfood <- (total_metabolites_food_inclusive/total_metabolites)*100
  
  results <- c(percent*100, perc_increase_G3, perc_increase_G4, perc_increase_allfood)
    return(results)
}

#how to call the function  
percent_annotation_overlap(G1_occurrence,G1G3_occurrence,G1G4_occurrence,G1G3G4_occurrence)

```

```{r}
#how to call the function for plasma  
percent_annotation_overlap(G2_occurrence,G2G3_occurrence,G2G4_occurrence,G2G3G4_occurrence)

```


#for datasets that have G1,G2, G3, G4; where G3 and G4 are both foods that should be taken into consideration
```{r}
library(dplyr)
#subsetting data for Q1 (increased knowledge because in a network with an annotated fecal metabolite)
G1_occurrence_libY <- G1_occurrence[G1_occurrence$LibraryID != "N/A", ]
G1_occurrence_libY_unique <-count(G1_occurrence_libY,componentindex) #creates unique list of component indices containing at least one annotation to compare our NAs to
G1_occurrence_libN <- G1_occurrence[G1_occurrence$LibraryID == "N/A", ]

G1_cluster_libY <- G1_occurrence_libY_unique %>%  #returns nodes that are NA that cluster with annotated nodes (includes self loops b/c script doesn't know they aren't connected all together)
  inner_join(G1_occurrence_libN, by='componentindex')

#remove single loops and do not include any overlaps between G1 and G3 or G4
a = subset(G1_cluster_libY, G1_cluster_libY$componentindex!= -1)
b = subset(a, a$G3==0)
a = subset(b, b$G4==0)
```


#subsetting the data for Q2a/b -- to increase knowledge because fecal sample clustering in a molecular family with a food metabolite (Q2a: unannotated) or (Q2b: annotated)
```{r}
G4_occurrence_no_sloops <- GNPS_network_data[GNPS_network_data$G4 !=0 | GNPS_network_data$G3 !=0, ]
G4_occurrence_no_sloops <- G4_occurrence_no_sloops[G4_occurrence_no_sloops$componentindex != -1, ]

G4_occurrence_no_sloops_unannotated <- G4_occurrence_no_sloops[G4_occurrence_no_sloops$LibraryID == 'N/A', ]
G4_occurrence_no_sloops_annotated <- G4_occurrence_no_sloops[G4_occurrence_no_sloops$LibraryID != 'N/A', ]

#Q2a input G4 to inner_join
G4_occurrence_no_sloops_unannotated_unique <-count(G4_occurrence_no_sloops_unannotated,componentindex) #creates unique list of component indices containing at least one annotation to compare our NAs to

#Q2b input G4 to inner_join
G4_occurrence_no_sloops_annotated_unique <-count(G4_occurrence_no_sloops_annotated,componentindex) #creates unique list of component indices containing at least one annotation to compare our NAs to

#Q2a/b input G1 to inner_join
G1_occurrence_no_sloops <- GNPS_network_data[GNPS_network_data$G1 != 0 & GNPS_network_data$componentindex != -1, ]

#inner_join to get the list of component_indeces where G1 and G3 or G4 are in the same molecular family; self loops already removed

Q2a <- G1_occurrence_no_sloops %>%
  inner_join(G4_occurrence_no_sloops_unannotated_unique, by='componentindex')

#will need to remove the ones present in G1 and G4
Q2a_no_G1G4 <- Q2a[Q2a$G3 == 0, ]
Q2a_no_G1G4 <- Q2a_no_G1G4[Q2a_no_G1G4$G4 == 0, ]

Q2b <- G1_occurrence_no_sloops %>%
  inner_join(G4_occurrence_no_sloops_annotated_unique, by='componentindex')

#will need to remove the ones present in G1 and G4
Q2b_no_G1G4 <- Q2b[Q2b$G3 == 0, ]
Q2b_no_G1G4 <- Q2b_no_G1G4[Q2b_no_G1G4$G4 == 0, ]

```

#script for calculating the percent increase in information gained from Q1 and Q2a/b - the total is from networking
Question 1: calculating percent annotation by the increase of unnannoted compounds found only in G1 that have new information because they network with an annotated compound -- found in G4 (could be an annotated G1G4 molecule or a G4 molecule)
Question 2: additional info because in the same cluster
```{r}

percent_annotation <- function(occurrence_table,new_annotations) {
  no_annot_cluster <- data.frame(dim(new_annotations))
  no_annot_cluster <- no_annot_cluster[1,1]
  
  dataframe_dim <- data.frame(dim(occurrence_table)) 
  total_metabolites <- dataframe_dim[1,1]                               
  percent <- no_annot_cluster / total_metabolites
  percent*100
}

#how to call the function  
Q1 <- percent_annotation(G1_occurrence,a)

Q2a <- percent_annotation(G1_occurrence,Q2a_no_G1G4)
Q2b <- percent_annotation(G1_occurrence,Q2b_no_G1G4)
#can't just do a straight addition - need to merge tables and take unique ones from 'cluster.index'

total <- rbind(a,Q2a_no_G1G4)
total2 <- rbind(total,Q2b_no_G1G4)
Total_added <-count(total2,cluster.index)
#creates unique list of cluster indices, yielding the total number of cluster with added info, regardless of via method Q1 or Q2a or Q2b
Q_added <- percent_annotation(G1_occurrence,Total_added)
  
```

=================================================================================================================

#For plasma analysis (human data in G2 of networking analysis)
```{r}
library(dplyr)
#subsetting data for Q1 (increased knowledge because in a network with an annotated plasma metabolite)
G2_occurrence_libY <- G2_occurrence[G2_occurrence$LibraryID != "N/A", ]
G2_occurrence_libY_unique <-count(G2_occurrence_libY,componentindex) #creates unique list of component indices containing at least one annotation to compare our NAs to
G2_occurrence_libN <- G2_occurrence[G1_occurrence$LibraryID == "N/A", ]
head(G2_occurrence_libY)
G2_cluster_libY <- G2_occurrence_libY_unique %>%  #returns nodes that are NA that cluster with annotated nodes (includes self loops b/c script doesn't know they aren't connected all together)
  inner_join(G2_occurrence_libN, by='componentindex')
head(G2_cluster_libY)
#remove single loops and do not include any overlaps between G2 and G3 or G4
a = subset(G2_cluster_libY, G2_cluster_libY$componentindex!= -1)
b = subset(a, a$G3==0)
a = subset(b, b$G4==0)
```

#subsetting the data for Q2a/b -- to increase knowledge because fecal sample clustering in a molecular family with a food metabolite (Q2a: unannotated) or (Q2b: annotated)
```{r}
G4_occurrence_no_sloops <- GNPS_network_data[GNPS_network_data$G4 !=0 | GNPS_network_data$G3 !=0, ]
G4_occurrence_no_sloops <- G4_occurrence_no_sloops[G4_occurrence_no_sloops$componentindex != -1, ]

G4_occurrence_no_sloops_unannotated <- G4_occurrence_no_sloops[G4_occurrence_no_sloops$LibraryID == 'N/A', ]
G4_occurrence_no_sloops_annotated <- G4_occurrence_no_sloops[G4_occurrence_no_sloops$LibraryID != 'N/A', ]

#Q2a input G4 to inner_join
G4_occurrence_no_sloops_unannotated_unique <-count(G4_occurrence_no_sloops_unannotated,componentindex) #creates unique list of component indices containing at least one annotation to compare our NAs to

#Q2b input G4 to inner_join
G4_occurrence_no_sloops_annotated_unique <-count(G4_occurrence_no_sloops_annotated,componentindex) #creates unique list of component indices containing at least one annotation to compare our NAs to

#Q2a/b input G1 to inner_join
G2_occurrence_no_sloops <- GNPS_network_data[GNPS_network_data$G2 != 0 & GNPS_network_data$componentindex != -1, ]

Q2a <- G2_occurrence_no_sloops %>%
  inner_join(G4_occurrence_no_sloops_unannotated_unique, by='componentindex')
head(Q2a)
#will need to remove the ones present in G1 and G4
Q2a_no_G2G4 <- Q2a[Q2a$G3 == 0, ]
Q2a_no_G2G4 <- Q2a_no_G2G4[Q2a_no_G2G4$G4 == 0, ]
head(Q2a_no_G2G4)

Q2b <- G2_occurrence_no_sloops %>%
  inner_join(G4_occurrence_no_sloops_annotated_unique, by='componentindex')
head(Q2b)
#will need to remove the ones present in G1 and G4
Q2b_no_G2G4 <- Q2b[Q2b$G3 == 0, ]
Q2b_no_G2G4 <- Q2b_no_G2G4[Q2b_no_G2G4$G4 == 0, ]
head(Q2b_no_G2G4)

```

#script for calculating the percent increase in information gained from Q1 and Q2a/b
Question 1: calculating percent annotation by the increase of unnannoted compounds found only in G1 that have new information because they network with an annotated compound -- found in G4 (could be an annotated G1G4 molecule or a G4 molecule)
a = subset(G1_cluster_libY, componentindex!= -1 & G4==0)
Question 2: additional info because in the same cluster
```{r}

percent_annotation <- function(occurrence_table,new_annotations) {
  no_annot_cluster <- data.frame(dim(new_annotations))
  no_annot_cluster <- no_annot_cluster[1,1]
  
  dataframe_dim <- data.frame(dim(occurrence_table)) 
  total_metabolites <- dataframe_dim[1,1]                               
  percent <- no_annot_cluster / total_metabolites
  percent*100
}

#how to call the function  
Q1 <- percent_annotation(G2_occurrence,a)

Q2a <- percent_annotation(G2_occurrence,Q2a_no_G2G4)
Q2b <- percent_annotation(G2_occurrence,Q2b_no_G2G4)
#can't just do a straight addition - need to merge tables and take unique ones from 'cluster.index'

total <- rbind(a,Q2a_no_G2G4)
total2 <- rbind(total,Q2b_no_G2G4)
Total_added <-count(total2,cluster.index)
#creates unique list of cluster indices, yielding the total number of cluster with added info, regardless of via method Q1 or Q2a or Q2b
Q_added <- percent_annotation(G2_occurrence,Total_added)
  
```

=================================================================================================================

Input data
```{r, echo=FALSE}
data_wTotal <-read.csv("../data/spectral_interpretation_rate.csv", header = TRUE) #, row.names= 1)  
```

Figure 2 generation
```{r}
#library("onewaytests")
library(rstatix)

#Overall----
data_human <- data_wTotal[data_wTotal$sample_type_broad == "Human" & data_wTotal$analysis == "yes", ]
human_welch <-welch.test(explanation_rate~annotation_type, data= data_human) #4.526236e-23 
games_howell_test(data_human, explanation_rate ~ annotation_type, conf.level = 0.95, detailed = FALSE)

get.bsbp_human <-function(data,factor,colorby){
  library(beeswarm) 
  BS <-beeswarm(data~factor,  pch=19, method="swarm", cex=2.0, xlab="",ylab="Interpreted Spectral Rate (%)", labels = c(""), pwcol=(colorby), corral="gutter" , par(cex.lab=1.2, cex.axis=1.5, cex.lab= 1.5, mar=c(5,5,5,2))) 
   BS
   boxplot <-boxplot(data~factor, add = T, col="#0000ff22", axes=FALSE, cex=2.0)
   boxplot
   legend('topleft', legend = c("CSF","Plasma", "Urine","Stool","Other"), col=c("magenta4", "violetred4", "palevioletred1","darkblue","purple4"), cex=1.35, pch=19,bty='n') 
  text(5.2, 2.5, paste("P =",signif(human_welch$p.value,2))
   )
}

human <- get.bsbp_human(data_human$explanation_rate,data_human$order,as.character(data_human$label_gen))

png("human_analysis_30June20.png", width = 8, height = 7.5, units = 'in', res = 300)
human_analysis <-get.bsbp_human(data_human$explanation_rate,data_human$order,as.character(data_human$label_gen))
dev.off()

#Fecal----
data_fecal <- data_wTotal[data_wTotal$sample_type == "Fecal" & data_wTotal$analysis == "yes", ]
fecal_welch <-welch.test(explanation_rate~annotation_type, data= data_fecal) #3.051825e-11 
games_howell_test(data_fecal, explanation_rate ~ annotation_type, conf.level = 0.95, detailed = FALSE)

get.bsbp_fecal <-function(data,factor,colorby){
  library(beeswarm) 
  BS <-beeswarm(data~factor,  pch=19, method="swarm", cex=2.0, xlab="",ylab="Interpreted Spectral Rate (%)", labels = c(""), pwcol=(colorby), corral="gutter" , par(cex.lab=1.2, cex.axis=1.5, cex.lab= 1.5, mar=c(5,5,5,2))) 
   BS
   boxplot <-boxplot(data~factor, add = T, col="#0000ff22", axes=FALSE, cex=2.0)
   boxplot
   legend('topleft', legend = c("Children with Medical Complexity","Malawi legume supplementation","Fermented food consumption","Rotarix vaccine response", "Impact of diet on RA","Sleep deprivation (n=2)","American Gut","Centenarian", "Infant", "IBD (n=6)"), col=c("darkslategray2","royalblue","steelblue2","dodgerblue4","deepskyblue","lightskyblue","blue","dodgerblue1","slategray2","darkblue"), cex=1.35, pch=19,bty='n') 
    text(5.2, 2.5, paste("P =",signif(fecal_welch$p.value,2)))
}

fecal <-get.bsbp_fecal(data_fecal$explanation_rate,data_fecal$order,as.character(data_fecal$label))

png("fecal_analysis_30June20_legend.png", width = 8, height = 7.5, units = 'in', res = 300)
fecal_analysis <-get.bsbp_fecal(data_fecal$explanation_rate,data_fecal$order,as.character(data_fecal$label))
dev.off()

#Plasma----
data_plasma <- data_wTotal[data_wTotal$sample_type_specific == "Plasma" & data_wTotal$analysis == "yes", ]
plasma_welch <- welch.test(explanation_rate~annotation_type, data= data_plasma) 
games_howell_test(data_plasma, explanation_rate ~ annotation_type, conf.level = 0.95, detailed = FALSE)


get.bsbp_plasma <-function(data,factor,colorby){
  library(beeswarm) 
  BS <-beeswarm(data~factor,  pch=19, method="swarm", cex=2.0, xlab="",ylab="Interpreted Spectral Rate (%)", labels = c(""), pwcol=(colorby), corral="gutter",par(cex.lab=1.2, cex.axis=1.5, cex.lab= 1.5, mar=c(5,5,5,2)) )
   BS
   boxplot <-boxplot(data~factor, add = T, col="#0000ff22", axes=FALSE, cex=2.0)
   boxplot
legend('topleft', legend = c("Impact of diet on RA", "Alzheimer's disease","Sleep deprivation (n=2)", "Centenarian",  "Covid-19"), col=c("maroon3","indianred1","violetred4","deeppink1","magenta"), pch=19, bty='n', cex= 1.35) 
    text(5.2, 3.5, paste("P =",signif(plasma_welch$p.value,2)))
}

plasma <- get.bsbp_plasma(data_plasma$explanation_rate,data_plasma$order,as.character(data_plasma$label))

png("plasma_analysis_30June20.png", width = 8, height = 7.5, units = 'in', res = 300)
plasma_analysis <-get.bsbp_plasma(data_plasma$explanation_rate,data_plasma$order,as.character(data_plasma$label))
dev.off()


#Other-----------
data_other <- data_wTotal[data_wTotal$sample_type_specific == "Human_other" & data_wTotal$analysis == "yes", ]
other_welch <- welch.test(explanation_rate~annotation_type, data= data_other) #5.67207e-05  
games_howell_test(data_other, explanation_rate ~ annotation_type, conf.level = 0.95, detailed = FALSE)

get.bsbp_humanother <-function(data,factor,colorby){
  library(beeswarm) 
  BS <-beeswarm(data~factor,  pch=19, method="swarm", cex=2.0, xlab="",ylab="Interpreted Spectral Rate (%)", labels = c(""), pwcol=(colorby), corral="gutter",par(cex.lab=1.2, cex.axis=1.5, cex.lab= 1.5, mar=c(5,5,5,2)) )
   BS
   boxplot <-boxplot(data~factor, add = T, col="#0000ff22", axes=FALSE, cex=2.0)
   boxplot
 legend('topleft', legend = c("Legume supplementation, urine","Alzheimer's disease, CSF","IBD_biopsy, tissue",  "Gout, serum", "Adult, saliva", "Infant, oral", "Infant, skin" ), col=c("palevioletred1","magenta4","darkorchid4","deeppink2","purple","purple4","darkmagenta"), pch=19, bty='n', cex= 1.35) 
    text(5.2, 5, paste("P =",signif(other_welch$p.value,2)))
}

other <-get.bsbp_humanother(data_other$explanation_rate,data_other$order,as.character(data_other$label))

png("other_analysis_30June20.png", width = 8, height = 7.5, units = 'in', res = 300)
not_plasma_analysis <-get.bsbp_humanother(data_other$explanation_rate,data_other$order,as.character(data_other$label))
dev.off()


#SSF----
data_CIAO <- data.frame(data_wTotal[data_wTotal$dataset_broad == "CIAO",])
data_ONR <- data_wTotal[data_wTotal$dataset_broad == "ONR",]
data_SSF <- rbind(data_CIAO,data_ONR)

data_Ciao.ONR_initial <- data_SSF[data_SSF$annotation_type == "initial" & data_SSF$analysis_SSF_test == "yes",]
data_Ciao.ONR_SSF <- data_SSF[data_SSF$annotation_type == "SSF" & data_SSF$analysis_SSF_test == "yes",]

D_Ciao.ONR_ItoSSF <- data_Ciao.ONR_SSF$explanation_rate - data_Ciao.ONR_initial$explanation_rate 
D_Ciao.ONR_ItoSSF <- data.frame(D_Ciao.ONR_ItoSSF)
D_Ciao.ONR_ItoSSF$dataset_broad <-data_Ciao.ONR_SSF$dataset_broad
D_Ciao.ONR_ItoSSF$sample_specific <-data_Ciao.ONR_SSF$sample_specific  
D_Ciao.ONR_ItoSSF$analysis <-data_Ciao.ONR_SSF$analysis 
D_Ciao.ONR_ItoSSF$label <- data_Ciao.ONR_SSF$label

SSF_ttest <- t.test(D_Ciao.ONR_ItoSSF$D_Ciao.ONR_ItoSSF ~ D_Ciao.ONR_ItoSSF$analysis, var.equal = FALSE) # p-value = 0.005124

get.bsbp_SSF <-function(data,factor,colorby){
  library(beeswarm) 
  BS <-beeswarm(data~factor,  pch=19, method="swarm", cex=2.0, xlab="",ylab="Interpreted Spectral Rate (%), Initial to SSF",  corral="gutter" ,labels=c("No", "Yes"), pwcol=  (colorby) ,par(cex.lab=1.2, cex.axis=1.5, cex.lab= 1.5, mar=c(5,5,5,2)))
   BS
   boxplot <-boxplot(data~factor, add = T, col="#0000ff22", axes=FALSE, cex=2.0)
   boxplot
    legend('topleft', legend = c( "Sleep deprivation, plasma", "Sleep deprivation, stool","Centenarian, plasma","Centenarian, stool") , col=c("violetred4","deepskyblue","deeppink3","dodgerblue1"), pch=19, bty='n', cex= 1.35)
    text(2.4, 4.5, paste("P =",signif(SSF_ttest$p.value,2)))
    }

SSF <- get.bsbp_SSF(D_Ciao.ONR_ItoSSF$D_Ciao.ONR_ItoSSF,D_Ciao.ONR_ItoSSF$analysis,as.character(D_Ciao.ONR_ItoSSF$label))

png("D_SSF_analysis_30June20.png", width = 8, height = 7.5, units = 'in', res = 300)
D_SSF_analysis <-get.bsbp_SSF(D_Ciao.ONR_ItoSSF$D_Ciao.ONR_ItoSSF,D_Ciao.ONR_ItoSSF$analysis,as.character(D_Ciao.ONR_ItoSSF$label))
dev.off()

```

Change in interpreted spectral rates
```{r}
library(ggplot2)  

data_fecal_initial <- data_fecal[data_fecal$annotation_type == "initial",]
data_fecal_total <-data_fecal[data_fecal$annotation_type == "TOTAL",]
data_fecal_GFOP <-data_fecal[data_fecal$annotation_type == "GFOP",]

D_annotation_fecal_ItoT <- data_fecal_total$explanation_rate - data_fecal_initial$explanation_rate
D_annotation_fecal_ItoG <- data_fecal_GFOP$explanation_rate - data_fecal_initial$explanation_rate

mean_se(D_annotation_fecal_ItoT)  #43.675 +/- 3.06078
mean_se(D_annotation_fecal_ItoG)  # 26.7968  +/- 3.29889

#------

data_plasma_initial<- data_plasma[data_plasma$annotation_type == "initial",]
data_plasma_total <-data_plasma[data_plasma$annotation_type == "TOTAL",]
data_plasma_GFOP<-data_plasma[data_plasma$annotation_type == "GFOP",]

D_annotation_plasma_ItoT <- data_plasma_total$explanation_rate - data_plasma_initial$explanation_rate
D_annotation_plasma_ItoG <- data_plasma_GFOP$explanation_rate - data_plasma_initial$explanation_rate

mean_se(D_annotation_plasma_ItoT)  #54.13741 +/- 4.51729
mean_se(D_annotation_plasma_ItoG)  # 27.64304  +/- 5.09078

#----

data_no_plasma_initial<- data_no_plasma[data_no_plasma$annotation_type == "initial",]
data_no_plasma_total <-data_no_plasma[data_no_plasma$annotation_type == "TOTAL",]
data_no_plasma_GFOP<-data_no_plasma[data_no_plasma$annotation_type == "GFOP",]

D_annotation_no_plasma_ItoT <- data_no_plasma_total$explanation_rate - data_no_plasma_initial$explanation_rate
D_annotation_no_plasma_ItoG <- data_no_plasma_GFOP$explanation_rate - data_no_plasma_initial$explanation_rate

mean_se(D_annotation_no_plasma_ItoT)  #57.99182	 +/- 4.16442
mean_se(D_annotation_no_plasma_ItoG)  # 40.55903  +/- 4.63956

#----

data_environment_initial<- data_environment[data_environment$annotation_type == "initial",]
data_environment_total <-data_environment[data_environment$annotation_type == "TOTAL",]
data_environment_GFOP<-data_environment[data_environment$annotation_type == "GFOP",]

D_annotation_environment_ItoT <- data_environment_total$explanation_rate - data_environment_initial$explanation_rate
D_annotation_environment_ItoG <- data_environment_GFOP$explanation_rate - data_environment_initial$explanation_rate

mean_se(D_annotation_environment_ItoT)  #26.4142 +/- 4.33048
mean_se(D_annotation_environment_ItoG)  # 10.51087  +/- 2.474996

```


